---


title: "BCB546_R_programming"
author: "Sristi"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Basic codes to load files into datastream, inspect them and merge two data files to process their data

```url_genotypes <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/fang_et_al_genotypes.txt"
print(url_genotypes)
url_snp_position <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/snp_position.txt"
genotypes <- read.table(url_genotypes, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
snp_position <- read.table(url_snp_position, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
head(genotypes)  # View the first few rows of genotypes data
head(genotypes)
head(snp_position)
str(genotypes)
str(snp_position)
dim(genotypes)
dim(snp_position)
object.size(genotypes)
object.size(snp_position)
num_lines_snp_position <- length(count.fields(url_snp_position, sep = "\t"))
num_lines_genotypes
url_transposed <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/transposed_genotypes.txt"
print(url_transposed)
transposed <- read.table(url_transposed, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
str(transposed)
str(snp_position) # Check the structure of the loaded data to ensure it's a data frame
url_transposed_genotypes <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/transposed_genotypes.txt"
transposed_genotypes <- read.table(url_transposed_genotypes, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
transposed_genotypes <- transposed_genotypes[-c(1, 2), ]
str(transposed_genotypes)
head(transposed_genotypes)
colnames(transposed_genotypes)
colnames(snp_position)
url_original_transposed_genotypes <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/transposed_genotypes.txt"
original_transposed_genotypes <- read.table(url_original_transposed_genotypes, header = TRUE, sep = "\t", stringsAsFactors = FALSE) #Reloaded the original transposed_genotypes file as a backup
head(original_transposed_genotypes)
url_transposed_raw <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/transposed_genotypes.txt" # URL for the transposed_genotypes.txt file on GitHub
url_transposed_raw <- read.table(url_transposed_raw, header = TRUE, sep = "\t", stringsAsFactors = FALSE) # Load the file from the GitHub URL
View(url_transposed_raw)
new_header <- as.character(url_transposed_raw[1, ]) # Extract the first row as a character vector
View(url_transposed_raw)
url_transposed_no_header <- url_transposed_raw[-1, ] # Remove the first row (old header)
colnames(url_transposed_no_header) <- new_header #Set the first row as column names
url_transposed_no_header_cleaned <- url_transposed_no_header[-1, ] #Remove the first row (which is now the header row)
head(url_transposed_no_header_cleaned) # View the cleaned data
new_header2 <- as.character(url_transposed_raw[2, ])
View(new_header2)
View(url_transposed_no_header)
url_transposed_no_header2 <- url_transposed_raw[-1, ]
colnames(url_transposed_no_header2) <- new_header2
url_transposed_no_header_cleaned2 <- url_transposed_no_header2[-1, ]
head(url_transposed_no_header_cleaned2) #This file has the groups as header and the rows that match with the snp_position file
colnames(url_transposed_no_header_cleaned2) #To check the current column names
colnames(url_transposed_no_header_cleaned2)[colnames(url_transposed_no_header_cleaned2) == "Group"] <- "SNP_ID" #To change the header name of the first column from "Oldname" to "Newname"
merged_data <- merge(url_snp_position, url_transposed_no_header_cleaned2, by = "SNP_ID", all = TRUE) #Merged the two data files url_snp_position and url_transposed_no_header_cleaned2
merged_data_backup <- merged_data #Created a backup for the merged data files (number of rows and columns got changed here, need to check that)
merged_data <- merged_data[, -2] #removed the column 2 which was cdv_marker that is not needed

###Pulled the columns that contains SNP_ID, Chromosome, Position, ZMMIL, ZMMLR, ZMMMR from the merged_data file to merged_data_subset file
selected_columns <- grep("SNP_ID|Chromosome|Position|ZMMIL|ZMMLR|ZMMMR", names(merged_data), value = TRUE) 
merged_data_subset <- merged_data[, selected_columns] 

#Inspect to confirm the samples in the file
count_headers_with_ZMMIL <- sum(grepl("ZMMIL", colnames(merged_data_subset))) #To count the number of headers that have ZMMIL
print(count_headers_with_ZMMIL) #[1] 290
count_headers_with_ZMMLR <- sum(grepl("ZMMLR", colnames(merged_data_subset))) #To count the number of headers that have ZMMLR
print(count_headers_with_ZMMLR) #[1] 1256
count_headers_with_ZMMMR <- sum(grepl("ZMMMR", colnames(merged_data_subset))) #To count the number of headers that have ZMMMR
print(count_headers_with_ZMMMR) #[1] 27

url_transposed_genotypes_version <- "https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2024/refs/heads/main/assignments/UNIX_Assignment/transposed_genotypes.txt"
transposed_genotypes_version <- read.table(url_transposed_genotypes_version, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
write.table(merged_data_subset, file = "merged_data_subset", sep = "\t", row.names = FALSE, quote = FALSE) #code to put files from data to working directory



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
